```{r global_options, include=FALSE}
library(knitr)
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE, fig.align="center", fig.height=4.5, fig.width=4.5)
```
---
output: html_document
---
## Homework 2
#### BIO5312
#### Due 9/12/17 by 5:30 pm

**Enter your name here**

<br>
This homework again uses the **sparrows.csv** dataset we used in homework 1. To begin, set your working directory (`Session -> Set Working Directory`) to the appropriate folder on your computer where the dataset is located, and read the dataset into R. You may also use the command `setwd()` if you are comfortable.

```{r}
sparrows <- read.csv("sparrows.csv")
```

For each question, show your code and its output (i.e. do not add either `echo=FALSE` or `results="hide"` to your R chunk settings!). For questions which ask you to provide a written response, please limit yourself to 1-3 concise sentences.

<br>

#### Question 1 (5 pts)
Repeat question 3c from homework 1 using `dplyr`: Determine the sex, age, and survival status for the bird with the longest and shortest wingspread. Complete this task in *1 line of code*, using the `%>%` operator to link your `dplyr` commands. *Hint: you will need to use the functions `filter` and `select`.*

```{r}
### Enter your code here.
```

<br>

#### Question 2 (5 pts).

Make a scatterplot of sparrow skull lengths (response variable) against sparrow skull widths (explanatory variable). Make two versions of this plot:
1. Set the point color based on wingspread
2. Set the point size based on wingspread. You may wish to include the `alpha` parameter when calling `geom_point()` for this plot.
Which version do you think is more informative?

```{r}
### Enter your code here.
```

**Provide your brief answer here.**

<br>

#### Question 2

<br>
**Question 2a (5 pts).** Use `dplyr` to summarize the mean and standard deviation of Tarsus lengths between Sex groupings (i.e. male and female). *Hint: your result should be a data frame with 2 rows.* 

```{r}
### Enter your code here.
```

<br>
**Question 2b (5 pts).** Now, use `dplyr` to summarize the mean and standard deviation of Tarsus lengths for all combinations of Sex and Age. Have the output printed arranged according to *descending* Tarsus length standard deviations. *Hint: your result should be a data frame with 4 rows.* 

```{r}
### Enter your code here.
```


<br>
**Question 2c (5 pts).** Use `ggplot2` to generate a line plot showing the differences between the mean Tarsus Length across sexes, showing separate lines for young vs adult birds. In other words, create a plot similar to this one: [example_2c.pdf](http://sjspielman.org/bio5312_fall2017/examples/example_2c.pdf): *Hint: you will need to specify a `group` aesthetic.* 

```{r}
### Enter your code here.
```


<br>
**Question 2d (5 pts).** Instead of merely showing the *means*, visualize the full distributions for the Tarsus Lengths across sexes and age. Make this plot for only the birds which survived. Make this figure two ways, using a violin plot and then using a boxplot. Which plot do you prefer for representing the data and why?  *Hint: you will _not_ need to specify a `group` aesthetic.* 

```{r}
### Enter your code here.

```

**Explain your plot preference here.**

<br>

#### Question 3

<br>
**Question 3a (5 pts).** Use `ggplot2` to make overlayed density plots (*hint: use the `alpha` parameter*) showing bird Weights for the Alive and Dead birds. 
```{r}
### Enter your code here.
```

<!--
<br>
**Question 3b (5 pts).** Use `dplyr` and `ggplot2` to make overlayed density plots (*hint: use the `alpha` parameter*) showing bird Weights for the Alive and Dead birds, specifically for young birds with a sternum length between 20--22 (including 20 and 22). 
```{r}
sparrows %>% filter(Sternum_Length >= 20, Sternum_Length <=22) %>% ggplot(aes(x = Weight, fill = Survival)) + geom_density(alpha = 0.4)
```
-->
<br>
**Question 3b (5 pts).** We are also interested in viewing "sparrow BMI" (defined for birds as weight divided by length). Use `dplyr` to create a new BMI column in the sparrows data, and then use `ggplot2` to make overlayed density plots of BMI for Alive vs Dead birds. Do you think "BMI" influenced survival? Why or why not?
```{r}
### Enter your code here.
```

**Explain why (not) BMI influenced survival here.**











