```{r global_options, include=FALSE}
#### Load any necessary libraries here. Some basics are loaded for you. ####
require(knitr)
require(dplyr)
require(tidyr)
require(ggplot2)
require(purrr)
require(readr)
require(broom)
knitr::opts_chunk$set(echo = TRUE, fig.align="center", fig.height=4.5, fig.width=4.5)
```
---
output: html_document
---

## Final Project (200 points)
### BIO5312

**Your name goes here.**

<br>

**Instructions can be found on Canvas and on the course website. Please read them carefully!! Additionally, consult the example project on the course website to aid your project preparation.**

<br><br>

## Introduction

Text for your introduction goes here. It should be formatted as "plain-text", as this text is written, with a bullet point describing variables. If references are used, please add a "References" section (two hashtag header) to the *end* of the document.

<br><br>

## Data Preparation

```{r}
### Any code used to prepare your data for analysis goes here.
### Include comments throughout this code explaining how you wrangled the data. Note that not every line needs a comment: Instead, each general task you perform (i.e. "extract from database") should have an associated comment.
## If no preparation is needed, simply read in your data here.
```

<br><br>

## Questions

<br>

#### Question 1

<br>

**To address this question, I will use _________:**

```{r}
### R code to perform statistical analysis goes here. 
```

Interpret your analysis here.

Intepret your figure here in a new paragraph.

```{r}
### R code visualize goes here.
```

<br>

#### Question 2

<br>

**To address this question, I will use _________:**


```{r}
### R code to perform statistical analysis goes here. 
```

Interpret your analysis here.

Intepret your figure here in a new paragraph.

```{r}
### R code visualize goes here.
```

<br>


#### Question 3 


<br>

**To address this question, I will use _________:**


```{r}
### R code to perform statistical analysis goes here. 
```

Interpret your analysis here.

Intepret your figure here in a new paragraph.

```{r}
### R code visualize goes here.
```

<br>

#### Question 4

<br>



**To address this question, I will use _________:**


```{r}
### R code to perform statistical analysis goes here. 
```

Interpret your analysis here.

Intepret your figure here in a new paragraph.

```{r}
### R code visualize goes here.
```

